---
title: "161.307: Assignment 2"
author: "John Lavack, 20001249"
format: 
  html:
    toc: true
embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r libraries, warning=FALSE}
#| label: setup
library(ggplot2)
library(dplyr)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(ggdist)
```

## Exercise 1

### Question 1: Which diet produces the heaviest chicks by day 21?

```{r}
#| label: ex1-q1

Q1_data <- as.data.frame(ChickWeight)

# your wrangling + chart code here
```




### Question 2: Which diet(s) have an unusual distribution of day-21 weights?

```{r}
#| label: ex1-q2

# your wrangling + chart code here
```

## Exercise 2

### Part 1: Purpose of the wrangling code

*(no more than 3 sentences)*

> YOUR ANSWER HERE

### Part 2: Reproduce the chart

```{r}
#| label: ex2-wrangling
data(highschool, package = "ggraph")

edge_weights <- highschool |>
  count(from, to, name = "weight")

g <- as_tbl_graph(edge_weights, directed = FALSE) |>
  activate(nodes) |>
  mutate(degree = centrality_degree())
```

```{r, eval = F}
#| label: ex2-chart
#| fig-width: 9
#| fig-height: 6

# modify this code to reproduce the target chart
# change the eval option of this chunk to execute it, once you 
# have put adequate values instead of the "??" below
set.seed(??)
ggraph(g, layout = "??") +
  geom_edge_link(aes(width = ??), alpha = ??) +
  geom_node_point(aes(colour = ??), size = ??) +
  theme_void()
```

## Exercise 3

```{r}
#| label: ex3-setup
# ---------------------------------------------------------------
# CHANGE THIS: replace 00 with the last two digits of your
# student ID, e.g. if your ID ends ...2107, use student_id <- 7
# ---------------------------------------------------------------
student_id <- 00
set.seed(student_id)

n <- sample(40:80, 1)
model_choice <- student_id %% 3

g <- switch(model_choice + 1,
  play_erdos_renyi(n, p = runif(1, 0.03, 0.08), directed = FALSE),
  play_barabasi_albert(n, power = runif(1, 0.8, 1.5), directed = FALSE),
  play_smallworld(n_dim = 1, dim_size = n, order = 3, p = runif(1, 0.01, 0.05))
) |>
  activate(nodes) |>
  mutate(degree = centrality_degree())

g
```

### Question

> YOUR QUESTION HERE

### Introduction

> YOUR INTRODUCTION HERE

### Approach

> YOUR APPROACH HERE

### Analysis

```{r}
#| label: ex3-wrangling

# resampling / simulation code here (at least two dplyr verbs)
```

```{r}
#| label: ex3-chart

# chart code here
```

### Discussion

> YOUR DISCUSSION HERE

## Reflection

1. *(Which of simulation/uncertainty/network thinking did you find hardest, and why?)* YOUR ANSWER HERE

2. *(An example where seeing the full distribution changed your interpretation, vs a "dynamite plot")* YOUR ANSWER HERE

3. *(Which Weeks 5-7 concept would you like to understand more deeply, and why?)* YOUR ANSWER HERE
