---
title: "161.307: Assignment 2"
author: "John Lavack, 20001249"
format: 
  html:
    toc: true
embed-resources: true
project:
  execute:
    # dir: document
    type: website
execute:
  dir: project
  echo: true
  warning: false
  message: false
---

```{r libraries, warning=FALSE, message = FALSE}
#| label: setup
library(ggplot2)
library(dplyr)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(ggdist)
#Uncomment and run if required for exercise 2, question 2. I didn't have Calibri set up. 
# library(showtext) 
# library(sysfonts)
```

```{r}
custom_palette <- c(
  "#a13939", 
  "#967348", 
  "#a59860", 
  "#5a903f", 
  "#285c50",
  "#18171e"
  )
```


## Exercise 1

For each of the following questions you are required to produce a chart that seeks to answer the question, focusing only on the final weigh-in (Time == 21).

Which diet produces the heaviest chicks by day 21, and how confident can we be in that ranking?

* Your chart must show a summary statistic (e.g. mean) for each diet together with an interval that communicates sampling uncertainty (e.g. SE or a confidence interval)
* Your chart must clearly state, in the labelling (axis title, subtitle or caption), exactly what the interval represents
* If you use colour you must choose a palette other than the defaults

### Question 1: Which diet produces the heaviest chicks by day 21?

```{r exercise 1, question 1}
#| label: ex1-q1
#| fig-width: 10
#| fig-height: 6

# To answer this I will do a ribbon plot. Grouping the data by diet.
# This will use geom_lineribbon from ggdist

Q1_data <- as.data.frame(ChickWeight)

Q1_data %>%
  # filter(Time == 21) %>%
  group_by(Diet, Time) %>%
  #create formula components for CIs
  summarise(
    n = n(),
    Mean_weight = mean(weight),
    se = sd(weight) / sqrt(n()),
    .groups = 'drop'
  ) %>%
  #calculate the CIs
  mutate(
    ci_50_lower = Mean_weight - qt(0.75, n - 1) * se,
    ci_50_upper = Mean_weight + qt(0.75, n - 1) * se,
    ci_80_lower = Mean_weight - qt(0.90, n - 1) * se,
    ci_80_upper = Mean_weight + qt(0.90, n - 1) * se,
    ci_95_lower = Mean_weight - qt(0.975, n - 1) * se,
    ci_95_upper = Mean_weight + qt(0.975, n - 1) * se,
  ) %>% 
  #pivoting to make it shaped like the data from mean_qi(), which was so easy to use
  pivot_longer(starts_with("ci_"), 
               names_to = c("type", ".width", "interval"),
               names_sep = "_",
               values_to = "value") %>%
  pivot_wider(names_from = "interval",
              names_prefix = ".",
              values_from = "value") %>%
  #making ci intervals factors for neat graph fill mapping
  mutate(.width = factor(sub("ci_", "", .width), 
                         levels = c("50", "80", "95"))) %>%
  #plotting, them geoms
  ggplot(aes(x= Time, y = Mean_weight)) +
  #this I found at https://mjskay.github.io/ggdist/articles/lineribbon.html
  geom_lineribbon(aes(ymin = .lower, ymax = .upper, fill = .width), 
                  col = '#a13939',
                  alpha = 0.4,
                  linewidth = 0.8) +
  facet_wrap(~ Diet, labeller = labeller(Diet = ~ paste("Diet", .))) +
  #Scale related commands
  # scale_color_manual(values = custom_palette) +
  #custom fill based off custom palette
  scale_fill_manual(values = c("95" = "#94AEA8", 
                               "80" = "#5E857C", 
                               "50" = "#285C50")) +
  #Labels
  labs(title = 'Confidence intervals for chick weight widen as chicks age',
       subtitle = 'As diets increase weight gain, their confidence intervals also widen',
       fill = 'Confidence Interval',
       x = "Time (Days since birth)",
       y = "Mean Chick Weight (grams)") +
  #themes
  theme_minimal() +
  theme(legend.position = 'top',
        plot.title = element_text(size = 15),
        strip.text = element_text(face = "bold"),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)))

#Tones for the fill
# "#94AEA8"
# "#5E857C"
# "#285C50"
# "#1E453C"
# "#142E28"

```


### Question 2: Which diet(s) have an unusual distribution of day-21 weights?

```{r exercise 1 graph 2}
#| label: ex1-q2
#| fig-width: 10
#| fig-height: 8


# I will answer this with a rain cloud plot from gg dist. 
#this will show the desnity on top with the point beneath

widths = c(0.5, 0.8, 0.95, 1)
mean_chick_weight <- Q1_data %>% filter(Time == 21) %>% summarise(x = round(mean(weight), 2))

#little data pret
Q1_data %>%
  filter(Time == 21) %>%
  #graph geoms
    #hdensity
  ggplot(aes(y = Diet, x = weight, fill = Diet)) +
  stat_halfeye(
    mapping = aes(fill_ramp = after_stat(level)),
    fill_type = "segments",
    .width = widths,
    show.legend = c(size = FALSE, fill_ramp = TRUE),
    col = NA,
    scale = 0.8 #shrinking to find best look without overlap
  ) +
    #make it rain!
  stat_dotsinterval(
    mapping = aes(fill_ramp = after_stat(level)),
    .width = widths,
    side = 'bottom',
    layout = 'weave',
    scale = 0.7,
    slab_linewidth = NA,
    col = NA,
    show.legend = FALSE,
    quantiles = 100 #each dot represents a part of a chick, made it like this to see the distribution unsmoothed better
  ) +
    #add a dot at the mean
  stat_summary(geom = "point", fun = mean, col = '#18171e', size = 3) +
    #add the vertical line for overall mean
  geom_vline(xintercept = mean_chick_weight$x, col = '#18171e', lty = 'dashed', linewidth = 1) +
    #annotation
  annotate("text",
           x = 225,
           y = 4.7, 
           label = paste0("Pooled mean = ", mean_chick_weight$x),
           size = 5, 
           hjust = -0.05
  ) +
  annotate("text",
         x = 150,
         y = 1, 
         label = "Diet 4 is right skewed. \nMany chicks are at the lightest extreme \nof the distribution \nand many at the mean.",
         size = 3
  ) +
  #scale commands
    #colour & legend
  scale_fill_manual(values = custom_palette) +
  scale_fill_ramp_discrete(
    from = "white",
    na.translate = FALSE
  ) +
  guides(fill = "none", fill_ramp = guide_legend(title = "interval")) +
    # axis
  scale_y_discrete(
    limits = rev(levels(Q1_data$Diet)),
    labels = ~paste0("Diet ", .)
    ) +
  scale_x_continuous(
    breaks = seq(
      from = min(Q1_data$weight),
      to = max(Q1_data$weight),
      by = 20)
      ) +
  #labels
  labs(
    title = 'Day 21 chick weights vary greatly by diet',
    subtitle = 'Diet 1 has the lightest chicks on average by day 21', 
    y = "",
    x = "Chick Weight (grams)"
  ) +
  #themes
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.text.x = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    legend.position = 'top'
  )
  
  
```


## Exercise 2

### Part 1: Purpose of the wrangling code

*(no more than 3 sentences)*

> YOUR ANSWER HERE


### Part 2: Reproduce the chart

```{r exercise 2 wrangling code}
#| label: ex2-wrangling
data(highschool, package = "ggraph")

edge_weights <- highschool |>
  count(from, to, name = "weight")

g <- as_tbl_graph(edge_weights, directed = FALSE) |>
  activate(nodes) |>
  mutate(degree = centrality_degree())
```



```{r adding calibri if requried, echo = FALSE}
# font_add("Calibri", "C:/Windows/Fonts/calibri.ttf")
# 
# font_add(
#   family   = "Calibri",
#   regular  = "calibri.ttf",
#   bold     = "calibrib.ttf",
#   italic   = "calibrii.ttf",
#   bolditalic = "calibriz.ttf"
# )
# showtext_auto()
```

```{r exercise 2 graph, eval = TRUE}
#| label: ex2-chart
#| fig-width: 9
#| fig-height: 6

# approach: trial and error! 
set.seed(161307)
ggraph(g, layout = "kk") +
  geom_edge_link(aes(width = weight), alpha = 0.5, color = "#676767") + #used a colour picker
  #control size of the lines
  scale_edge_width(range = c(0.3, 2)) +
  geom_node_point(aes(colour = degree), size = 4) +
  #Scale commands
  scale_colour_viridis_c() +
  #labels
  labs(
    title = "Highschool friendship network: edge width = years friendship reported",
    subtitle = "Node colour = degree (number of distinct friends nominated across both years)",
    colour = "Node\ndegree",
    edge_width = "Years\nfriendship\nreported",
  ) +
    #guides ordering
  guides(
    colour     = guide_colorbar(order = 1),  
    edge_width = guide_legend(order = 2)  
  ) +
  # themes
    #Font looks like calibri?. 
  theme_graph(base_family = "Calibri") +
  theme(
    legend.position = "bottom",
    plot.margin = margin(10,10,10,10),
    plot.title = element_text(size = 28, face = 'plain'),
    plot.subtitle = element_text(size = 18, margin = margin(t=0, b=15))
  )
```



## Exercise 3

```{r}
#| label: ex3-setup
# ---------------------------------------------------------------
# CHANGE THIS: replace 00 with the last two digits of your
# student ID, e.g. if your ID ends ...2107, use student_id <- 7
# ---------------------------------------------------------------
student_id <- 00
set.seed(student_id)

n <- sample(40:80, 1)
model_choice <- student_id %% 3

g <- switch(model_choice + 1,
  play_erdos_renyi(n, p = runif(1, 0.03, 0.08), directed = FALSE),
  play_barabasi_albert(n, power = runif(1, 0.8, 1.5), directed = FALSE),
  play_smallworld(n_dim = 1, dim_size = n, order = 3, p = runif(1, 0.01, 0.05))
) |>
  activate(nodes) |>
  mutate(degree = centrality_degree())

g
```

### Question

> YOUR QUESTION HERE

### Introduction

> YOUR INTRODUCTION HERE

### Approach

> YOUR APPROACH HERE

### Analysis

```{r}
#| label: ex3-wrangling

# resampling / simulation code here (at least two dplyr verbs)
```

```{r}
#| label: ex3-chart

# chart code here
```

### Discussion

> YOUR DISCUSSION HERE

## Reflection

1. *(Which of simulation/uncertainty/network thinking did you find hardest, and why?)* YOUR ANSWER HERE

2. *(An example where seeing the full distribution changed your interpretation, vs a "dynamite plot")* YOUR ANSWER HERE

3. *(Which Weeks 5-7 concept would you like to understand more deeply, and why?)* YOUR ANSWER HERE
