---
title: "161.307: Assignment 2"
author: "John Lavack, 20001249"
format: 
  html:
    toc: true
embed-resources: true
project:
  execute:
    # dir: document
    type: website
execute:
  dir: project
  echo: true
  warning: false
  message: false
---

```{r libraries, warning=FALSE, message = FALSE}
#| label: setup
library(ggplot2)
library(dplyr)
library(tidygraph)
library(ggraph)
library(tidyverse)
library(ggdist)
library(distributional)
library(igraph)
library(patchwork)
#Uncomment and run if required for exercise 2, question 2. I didn't have Calibri set up. 
library(showtext)
library(sysfonts)
```

```{r}
custom_palette <- c(
  "#a13939", 
  "#967348", 
  "#a59860", 
  "#5a903f", 
  "#285c50"
  # "#18171e"
  )
```


## Exercise 1

For each of the following questions you are required to produce a chart that seeks to answer the question, focusing only on the final weigh-in (Time == 21).

Which diet produces the heaviest chicks by day 21, and how confident can we be in that ranking?

* Your chart must show a summary statistic (e.g. mean) for each diet together with an interval that communicates sampling uncertainty (e.g. SE or a confidence interval)
* Your chart must clearly state, in the labelling (axis title, subtitle or caption), exactly what the interval represents
* If you use colour you must choose a palette other than the defaults

### Question 1: Which diet produces the heaviest chicks by day 21?

```{r exercise1 graph 1}
#| fig-width: 10
#| fig-height: 6


# wrangling
Q1_data_CI <- Q1_data %>% 
  group_by(Diet) %>%
  summarise(df = n() -1,
            n = n(), 
            mu = mean(weight),
            sigma = sd(weight) / sqrt(n())
            ) 

Q1_data_CI %>%
  #Plotting geoms
  ggplot(aes(y = Diet, 
             #small sample size, use t dist, gives a bit wider tails, less certainty
             xdist = dist_student_t(
               df = df, mu = mu, sigma = sigma))
         ) +
  stat_interval(
    mapping = aes(fill = after_stat(level)),
    .width = c(0.50, 0.80, 0.95, 0.99),
    show_point = TRUE
  ) +
  #Scale commands
  scale_color_manual(values = custom_palette) +
  # guides(colour = 'none') +
  scale_y_discrete(
    limits = rev(levels(Q1_data$Diet)),
    labels = ~paste0("Diet ", ., ' (n =', Q1_data_CI$n,')')
  ) +
  scale_x_continuous(
    breaks = seq(
      from = 120,
      to = 360,
      by = 20)
  ) +
  #Labels
  labs(title = 'Diet 3 has highest mean chick weight gain 21 days after hatching',
     subtitle = "It's wide confidence interval implies high variability, across the low sample size",
     fill = 'Confidence Interval',
     y = "",
     x = "Mean Chick Weight (grams)",
     colour = "Confidence Interval") +
  #themes
  theme_minimal() +
  theme(legend.position = 'top',
        # legend.title = 'Confidence Interval',
        plot.title = element_text(size = 15),
        axis.title.x = element_text(margin = margin(t = 10)),
        axis.title.y = element_text(margin = margin(r = 10)),
        panel.grid.major.y = element_blank()
  )

```

### Question 2: Which diet(s) have an unusual distribution of day-21 weights?

```{r exercise 1 graph 2}
#| label: ex1-q2
#| fig-width: 10
#| fig-height: 8


# I will answer this with a rain cloud plot from gg dist. 
#this will show the desnity on top with the point beneath

widths = c(0.5, 0.8, 0.95, 1)
mean_chick_weight <- Q1_data %>% filter(Time == 21) %>% summarise(x = round(mean(weight), 2))

#little data pret
Q1_data %>%
  filter(Time == 21) %>%
  #graph geoms
    #hdensity
  ggplot(aes(y = Diet, x = weight, fill = Diet)) +
  stat_halfeye(
    mapping = aes(fill_ramp = after_stat(level)),
    fill_type = "segments",
    .width = widths,
    show.legend = c(size = FALSE, fill_ramp = TRUE),
    col = NA,
    scale = 0.8 #shrinking to find best look without overlap
  ) +
    #make it rain!
  stat_dotsinterval(
    mapping = aes(fill_ramp = after_stat(level)),
    .width = widths,
    side = 'bottom',
    layout = 'weave',
    scale = 0.7,
    slab_linewidth = NA,
    col = NA,
    show.legend = FALSE,
    quantiles = 100 #each dot represents a part of a chick, made it like this to see the distribution unsmoothed better
  ) +
    #add a dot at the mean
  stat_summary(geom = "point", fun = mean, col = '#18171e', size = 3) +
    #add the vertical line for overall mean
  geom_vline(xintercept = mean_chick_weight$x, col = '#18171e', lty = 'dashed', linewidth = 1) +
    #annotation
  annotate("text",
           x = 225,
           y = 4.8, 
           label = paste0("Pooled mean = ", mean_chick_weight$x),
           size = 5, 
           hjust = -0.05
  ) +
  annotate("text",
         x = 325,
         y = 4.3, 
         label = "Diet 1's multimodal distribution\n is hidden by the density curve",
         size = 3
  ) +
  annotate("text",
       x = 325,
       y = 3.4, 
       label = "Diet 2's mean falls between\n two peaks, indicating two clusters",
       size = 3
  ) +
    annotate("text",
         x = 100,
         y = 2, 
         label = "Diet 3's left skew shows a\n high level of heavier chicks",
         size = 3
  ) +
  annotate("text",
         x = 150,
         y = 1, 
         label = "Diet 4 is right skewed. \nMany chicks are at the lightest extreme \nof the distribution \nand many at the mean.",
         size = 3
  ) +
  #scale commands
    #colour & legend
  scale_fill_manual(values = custom_palette) +
  scale_fill_ramp_discrete(
    from = "white",
    na.translate = FALSE
  ) +
  guides(fill = "none", fill_ramp = guide_legend(title = "Confidence interval")) +
    # axis
  scale_y_discrete(
    limits = rev(levels(Q1_data$Diet)),
    labels = ~paste0("Diet ", .)
  ) +
  scale_x_continuous(
    breaks = seq(
      from = min(Q1_data$weight),
      to = max(Q1_data$weight),
      by = 20)
  ) +
  #labels
  labs(
    title = 'Day 21 chick weights vary greatly by diet',
    subtitle = 'Diet 1 has the lightest chicks on average by day 21, Diet 4, the heaviest', 
    y = "",
    x = "Chick Weight (grams)"
  ) +
  #themes
  theme_minimal() +
  theme(
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.text.x = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    legend.position = 'top'
  )
  
  
```


## Exercise 2

### Part 1: Purpose of the wrangling code

*(no more than 3 sentences)*

First, the raw graph data is wrangled to have one row per unique edge using `count()` (aka single weighted edge pair), which is effectively `group_by(to, from) %>% summarise(weight = n())` and shows us if a student has a friend in one or both years (weight 1 and weight 2 respectively).  
Next the data is converted to an object tidygraph can use, which treats relationships as symmetric (not directed), so we are assuming friendship is mutual.  
Lastly it creates degree as the number of connections each node has using `centrality_degree()`, which gives the ability to visualise how each student (node) has rated their friends over the two years of interest.  

### Part 2: Reproduce the chart

```{r exercise 2 wrangling code}
#| label: ex2-wrangling
data(highschool, package = "ggraph")

edge_weights <- highschool |>
  count(from, to, name = "weight")

g <- as_tbl_graph(edge_weights, directed = FALSE) |>
  activate(nodes) |>
  mutate(degree = centrality_degree())
```

```{r adding calibri if requried, echo = FALSE}
font_add("Calibri", "C:/Windows/Fonts/calibri.ttf")

font_add(
  family   = "Calibri",
  regular  = "calibri.ttf",
  bold     = "calibrib.ttf",
  italic   = "calibrii.ttf",
  bolditalic = "calibriz.ttf"
)
showtext_auto()
```

```{r exercise 2 graph, eval = TRUE}
#| label: ex2-chart
#| fig-width: 9
#| fig-height: 6

# approach: trial and error! 
set.seed(161307)
ggraph(g, layout = "kk") +
  geom_edge_link(aes(width = weight), alpha = 0.5, color = "#676767") + #used a colour picker
  #control size of the lines
  scale_edge_width(range = c(0.3, 2)) +
  geom_node_point(aes(colour = degree), size = 4) +
  #Scale commands
  scale_colour_viridis_c() +
  #labels
  labs(
    title = "Highschool friendship network: edge width = years friendship reported",
    subtitle = "Node colour = degree (number of distinct friends nominated across both years)",
    colour = "Node\ndegree",
    edge_width = "Years\nfriendship\nreported",
  ) +
    #guides ordering
  guides(
    colour = guide_colorbar(order = 1),  
    edge_width = guide_legend(order = 2)  
  ) +
  # themes
    #Font looks like calibri?. 
  theme_graph(base_family = "Calibri") +
  theme(
    legend.position = "bottom",
    plot.margin = margin(10,10,10,10),
    plot.title = element_text(size = 28, face = 'plain'),
    plot.subtitle = element_text(size = 18, margin = margin(t=0, b=15))
  )
```



## Exercise 3

```{r}
#| label: ex3-setup

student_id <- 49
set.seed(student_id)

n <- sample(40:80, 1) #7-
model_choice <- student_id %% 3

g <- switch(model_choice + 1,
  play_erdos_renyi(n, p = runif(1, 0.03, 0.08), directed = FALSE),
  play_barabasi_albert(n, power = runif(1, 0.8, 1.5), directed = FALSE), #this is my one
  play_smallworld(n_dim = 1, dim_size = n, order = 3, p = runif(1, 0.01, 0.05))
) |>
  activate(nodes) |>
  mutate(degree = centrality_degree(),
         betweenness = centrality_betweenness(directed = FALSE),
         ) |>
  activate(edges) |>
  mutate(edge_betweenness = centrality_edge_betweenness(directed = FALSE))

```

```{r}
#| fig-width: 9
#| fig-height: 6

#stress and KK were similar in look, but kk is the easier one to read.
#try directed, try weights on edges. 
ggraph(g, layout = "kk") +
  geom_edge_link(aes(width = edge_betweenness), alpha = 0.4, color = "#676767") + 
  scale_edge_width(range = c(0.5, 4),
                   name = 'Edge\nbetweenness') +
  geom_node_point(aes(colour = betweenness, size = degree)) +
  #Scale commands
  scale_colour_gradientn(colours = rev(custom_palette)) +
  #labels
  labs(
    title = "Preferential attachment graph",
    subtitle = paste0("Each node represents an individual, each edge, contact between two.\nAverage betweenness = ", 
                      round(mean(betweenness(g, directed = FALSE)),2)),
    size = "Node\ndegree",
    colour = "Node\nbetweenness"
  ) +
  # guides ordering
  guides(
    colour = guide_colorbar(order = 1),
    size = guide_legend(order = 2),
    width = guide_legend(order = 3)
  ) +
  # themes
  theme_graph(base_family = "Calibri") +
  theme(
    legend.position = "bottom",
    plot.margin = margin(10,10,10,10),
    plot.title = element_text(size = 28, face = 'plain'),
    plot.subtitle = element_text(size = 12, margin = margin(t=0, b=15))
  )

```


### Question

How does average betweenness of this graph's node attachment vary under resampling?. 

### Introduction

> YOUR INTRODUCTION HERE

### Approach

> YOUR APPROACH HERE

Note to self: betweenness is a measure of traffic of shortest paths passing through the node. 
How many shortest paths pass through the node


To analyse the summary statistic parametric resampling will be applied - bootstrapping using the same parameters. 
This is the appropriate choice because with a Barabasi-Albert graph, the final graph is the end result of a stochastic process which grows stepwise through time.
A process of preferential attachment. Resampling the same graph/nodes with replacement would be to undo the preferential attachment structure. The uncertainty of interest is the model uncertainty, i.e. how much the mean betweenness varies across the same Barabasi-Albert process, not the betweenness from the observed final dataset.
Because the resampling process used is frequentist, the uncertainty is best shown with confidence intervals. 

### Analysis

```{r}
#| label: ex3-wrangling

#Replicate the graph process B time for a graph with n nodes. 
#Keep all the graph data, calculate the statistic for betweenness. Return as tibble.
#use this to visualise charts to show structure of the BA graph.

Q3_data <- function(n, B = 1000) {
  lapply(seq_len(B), function(i) {
    #create graph process
    graph_model <- play_barabasi_albert(n, power = runif(1, 0.8, 1.5), directed = FALSE)   
    
    graph_model %>%
      activate(nodes) %>%
      mutate(node_id = 1:n, #get node ID for visualisation
             betweenness = centrality_betweenness(directed = FALSE),
             replication = i) %>% 
      as_tibble() %>%
      select(replication, node_id, betweenness)
  }) %>%
    bind_rows()
}

#create the dataset, using n from the graph setup
Q3_graph_data <- Q3_data(n, B = 1000)

#wrangle data set for the ribbon
ribbon_data <- Q3_graph_data %>%
  group_by(node_id) %>%
  reframe( #gor a warning using summarise
    mean   = mean(betweenness),
    .width = c(0.5, 0.8, 0.95),
    .lower = c(quantile(betweenness, 0.25),
               quantile(betweenness, 0.10),
               quantile(betweenness, 0.025)),
    .upper = c(quantile(betweenness, 0.75),
               quantile(betweenness, 0.90),
               quantile(betweenness, 0.975)),
    .groups = 'drop'
  ) %>%
  mutate(.width = factor(.width, levels = c(0.5, 0.8, 0.95)))
  
```


```{r exercise 3 chart}
#| label: ex3-chart
#| fig-width: 9
#| fig-height: 6

#I want to look at the summary stat as per hte question, but I also want to see
#how the summary stat varies structurally. BA graphs are "ordered" you could say.
#Nodes are sequentially attached to the graph, so the noe number is a proxy for 
#how long it has been in the graph, and therefore the node's betweenness will
#likely reflect that. 

#===== Rain cloud chart =====#
widths = c(0.5, 0.8, 0.95)

#set up geoms
rain <- Q3_graph_data %>%
  group_by(replication) %>%
  summarise(betweenness = mean(betweenness)) %>%
  #graph geoms
    #hdensity
  ggplot(aes(x = 0, y = betweenness)) +
  stat_halfeye(
    mapping = aes(fill_ramp = after_stat(level)),
    fill_type = "segments",
    .width = widths,
    show.legend = c(size = FALSE, fill_ramp = TRUE),
    col = NA,
    fill = "#5a903f",
  ) +
    #make it rain!
  stat_dotsinterval(
    mapping = aes(fill_ramp = after_stat(level)),
    .width = widths,
    side = 'bottom',
    layout = 'bin',
    slab_linewidth = NA,
    col = NA,
    fill = "#5a903f",
    show.legend = FALSE,
    quantiles = 200 #each dot is 5 units
  ) +
    #add a dot at the mean
  stat_summary(geom = "point", fun = mean, col = '#18171e', size = 3) +
  #scale commands
    #colour & legend
  scale_fill_ramp_discrete(
    from = "white",
    na.translate = FALSE
  ) +
  guides(fill = "none", fill_ramp = guide_legend(title = "Confidence interval")) +
    # axis
  scale_y_continuous(
    breaks = seq(
      from = 40,
      to = 200,
      by = 20),
    limits = c(40, 220)
  ) +
  #labels
  labs(
    title = '',
    subtitle = 'As replications grow,\nthe distribution of mean betweenness\napproximates a normal distribution.', 
    y = "Mean Betweenness",
    x = ""
  ) +
  #themes
  theme_minimal(base_family = "Calibri") +
  theme(
    plot.subtitle = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 12, margin = margin(b = 10)),
    axis.text.x = element_blank(),
    legend.position = 'top',
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),   
  )


#===== Ribbon Chart =====#
#set up geoms
ribbon <- ggplot(ribbon_data, aes(x = node_id, y = mean, ymin = .lower, ymax = .upper)) +
  geom_lineribbon(aes(fill = .width), 
                  alpha = 0.4,
                  colour = NA) +
  geom_line(aes(y= mean, colour = "Mean betweenness"), linewidth = 1) +
  #Scale commands
  scale_fill_manual(values = c('0.5' = "#5A903F", 
                               '0.8' = "#9CBC8C", 
                               '0.95' = "#DEE9D9"),
                    guide = 'none') + #will use the rain cloud legend
  scale_colour_manual(values = c("Mean betweenness" = "#a13939"), name = NULL) +

  #axis
  scale_x_continuous(
    breaks = seq(
    from = 0,
    to = max(Q3_graph_data$node_id),
    by = 5),
    expand = expansion(mult = 0, add = 0)
  ) +
  scale_y_continuous(
    breaks = seq(
    from = 0,
    to = max(Q3_graph_data$betweenness),
    by = 100),
    expand = expansion(mult = 0, add = 0)
  ) +
  #labels
  labs(
    title = "",
    subtitle = "As new individuals interact, fewer paths traverse them,\nbetweenness tends to zero, and the confidence interval narrows\n towards zero.",
    x = "Node (attachment order)", 
    y = "Betweenness",
    ) +
  #theme
  theme_minimal(base_family = 'Calibri') +
  theme(
    legend.position = 'top',
    legend.justification = 'left',
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 15),    
  )

#===== Patchwork Output=====#
rain + ribbon + 
  plot_layout(ncol = 2, widths = c(0.3, 0.7)) +
  plot_annotation(title = "Preferential attachment graph node betweenness from 1000 replications") &
  theme(plot.title = element_text(family = "Calibri", size = 25))
  
```


### Discussion

> YOUR DISCUSSION HERE

## Reflection

1. *(Which of simulation/uncertainty/network thinking did you find hardest, and why?)* YOUR ANSWER HERE

2. *(An example where seeing the full distribution changed your interpretation, vs a "dynamite plot")* YOUR ANSWER HERE

3. *(Which Weeks 5-7 concept would you like to understand more deeply, and why?)* YOUR ANSWER HERE
